brados: declarative, programmable object storage

### 2016-05-25 paper.tex
paper update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

%%% why is it hard / why do naive approaches fail
In addition to the growth in the quantity of operations in use throughout Ceph
installations, Figure~\ref{fig:objclass-dev} also depicts the amount of
low-level C++ written to implement these operations. Unfortunately, this code
is written assuming a performance profile defined by the combination of the
hardware and software versions available at the time of development.  While
the bulk of these interfaces are created by core Ceph developers with a
complete view of the performance model, this may be changing as the
development community has been receptive to outside contributions with the
recent inclusion by CERN developers of an extension for performing limited
numeric operations on object data~\cite{cls_numops}. And while Ceph has not
yet reached the point of directly exposing these features to
non-administrative users, the under review inclusion of a mechanism for dynamically
defining extensions using Lua~\cite{cls_lua} suggests that aspects of this
feature may soon appear. What is needed is support for creating storage
interfaces using a method that allows transparent optimization as the system,
application, and supporting environment evolve.

% why hasn't it been solved before / why have previous approaches failed
Unfortunately the size of the design space for even simple interfaces can
be very large, and it can be difficult to choose a design that future proofs an
implementation against future hardware and software upgrades.
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the former is
primarily concerned with fixed interfaces, application-specific interfaces
built using active storage techniques have not to the best of our knoweldge
addressed portability concerne that arise from defining interfaces
imperatively. We address this gap by using a declarative language we call
\emph{brados} to define application-specific storage interfaces in
object-based storage systems, decoupling interface definition from
implementation and allowing interfaces to adapt to system changes without
modification.

To demonstrate the use of the \emph{brados} language we will show how a
high-performance shared-log based on the CORFU protocol can be built, as well
as existing real-world interfaces used in the Ceph storage system. First we
discuss the Ceph storage system which our prototype is built upon and the 
motivating system, CORFU. We then present the challenges that programmers face
when navigating the design space during the process of building an
application-storage interface. The \emph{brados} language is then described
using the CORFU system and an existing Ceph-specific interface as examples.
We conclude by discussing the optimization opportunities that we can realize
using analysis of interfaces. I hope you enjoy the show.

\section{Background and Motivation}

In this section we highlight the salient components of Ceph, especially its
\emph{object class} feature that offers users the ability to load and execute
application-specific codes. We provide a description of our motivating
example, a high-performance distributed shared-log built upon Ceph that makes
extensive use of the object class facility.

\subsection{Ceph and Storage Programmability}
\label{sec:objclass}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{.48\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{figures/ceph}
      \caption{Ceph Cluster}
      \label{fig:ceph}
  \end{subfigure}\quad
  \begin{subfigure}[b]{.40\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{figures/osd}
      \caption{OSD Arch}
      \label{fig:osd}
  \end{subfigure}
  \caption{The Ceph cluster is composed of object storage devices (OSD) that
  support object transactions and a rich set of hardware and software
configurations.}
\end{figure}

@@

The native object operations in RADOS roughly fall into two categories based
on the type of data being accessed: key-value items, or bulk bytestream data.
The key-value interface operates as a dedicated database associated with each
object, and the bytestream interface supports random byte-level access similar
to a file.  At a
low-level each of these abstract I/O interfaces map to hardware storage
devices through a pluggable object backend storage service. For instance,
LevelDB or RocksDB may be used to store key-value data, while the
\emph{FileStore} implementation maps the bytestream interface onto a local
POSIX file system~\cite{leveldb,rocksdb}. Several backend implementations
exist for storing data in targets such as the Kinetic Drive or in NVMe devices,
among others.

>>>>

%%% why is it hard / why do naive approaches fail
In addition to the growth in the quantity of operations in use throughout Ceph
installations, Figure~\ref{fig:objclass-dev} also depicts the amount of
low-level C++ written to implement these operations. Unfortunately, this code
is written assuming a performance profile defined by the combination of the
hardware and software versions available at the time of development.  While
the bulk of these interfaces are created by core Ceph developers with a
complete view of the performance model, this may be changing as the
development community has been receptive to outside contributions with the
recent inclusion by CERN developers of an extension for performing limited
numeric operations on object data~\cite{cls_numops}. And while Ceph has not
yet reached the point of directly exposing these features to
non-administrative users, the inclusion of a mechanism for dynamically
defining extensions using Lua~\cite{cls_lua} (currently pending review)
suggests that aspects of this
feature may soon appear. What is needed is support for creating storage
interfaces using a method that allows transparent optimization as the system,
application, and supporting environment evolve.

% why hasn't it been solved before / why have previous approaches failed
Unfortunately the size of the design space for even simple interfaces can
be very large, and it can be difficult to choose a design that future proofs an
implementation against hardware and software upgrades.
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the former is
primarily concerned with fixed interfaces, application-specific interfaces
built using active storage techniques have not to the best of our knoweldge
addressed portability concerns that arise from defining interfaces
imperatively. We address this gap by using a declarative language we call
\emph{brados} to define application-specific storage interfaces in
object-based storage systems, decoupling interface definition from
implementation and allowing interfaces to adapt to system changes without
modification.

To demonstrate the use of the \emph{brados} language we will show how a
high-performance shared-log based on the CORFU protocol can be built in Ceph, as well
as existing real-world interfaces used in the Ceph storage system. First we
discuss the Ceph storage system which our prototype is built upon and the 
motivating system, CORFU. We then present the challenges that programmers face
when navigating the design space during the process of building an
application-storage interface. The \emph{brados} language is then described
using the CORFU system and an existing Ceph-specific interface as motivating examples.
We conclude by discussing the optimization opportunities that we can realize
using analysis of interfaces. We hope you enjoy the show.

\section{Background}

In this section we describe the salient components of Ceph that we use to
construct application-specific extensions, and provide an overview of the
CORFU distributed shared-log which is the primary motivating example used in
this paper. The content of this section sets the stage for a discussion of the
complexities developers face building application-specific storage interfaces.

\subsection{Ceph Basics and Storage Interfaces}
\label{sec:objclass}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{.48\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{figures/ceph}
      \caption{Ceph Cluster}
      \label{fig:ceph}
  \end{subfigure}\quad
  \begin{subfigure}[b]{.40\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{figures/osd}
      \caption{OSD}
      \label{fig:osd}
  \end{subfigure}
  \caption{The Ceph cluster is composed of object storage devices (OSD) that
  provide access to objects on behalf of client requests. An OSD supports a
    wide variety of hardware and software configurations, as well as the
    transactional composition of object operations for defining
    application-level interfaces.}
\end{figure}

@@

The native object operations in RADOS roughly fall into two categories based
on the type of data being accessed: key-value items, or bulk bytestream data.
The key-value interface operates as an isolated database associated with each
object, and the bytestream interface supports random byte-level access similar
to a standard file interface.  At a low-level each of these abstract I/O
interfaces map to hardware storage devices through a pluggable object backend
shown in Figure~\ref{fig:osd}. For instance, LevelDB or RocksDB may be used to
store key-value data, while the \emph{FileStore} implementation maps the
bytestream interface onto a local POSIX file system~\cite{leveldb,rocksdb} via
XFS or other supported file systems. Several backend implementations exist for
storing data in a range of targets such as HDD, SSD, as well as
Ethernet-attached disks, and NVMe devices using optimized access libraries.
%and in SMR disks using a specialized backend called BlueStore.
